### 2024-02-24 products/manuscript/Manuscript.qmd

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../assets/dataanalysis-references.bib
csl: ../assets/apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract
_Write a summary of your project._

{{< pagebreak >}}

# Introduction

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

Finite mixture models are used to analyze data presumably arising from several separate, yet unobserved, groups. Among these models, Gaussian mixture models and mixture of t-distributions are particularly prevalent. However, such models which are based on symmetric distribution often do not accurately represent data with asymmetric distribution. Skewed data can be more effectively managed using asymmetric distributions, thereby extending their applicability across a broader spectrum of model-based clustering scenarios [1]. If the observed data is positive, distribution is right-skewed and assumed to come from a mixture distribution, employing mixture of gamma density is a rational choice [2].

This study intends to identify distinct customer groups within sales data, with each group comprising customers who exhibit similar characteristics in terms of purchase frequency and amounts. Modeling customer spending as a mixture of gamma distributions helps to identify different spending behaviors. The objective is to develop an Expectation Maximization (EM) algorithm which segregates groups based on their spending patterns, considering the individual purchases transactions made by each customer.

The exploration begins with applying commonly used alternative models to a publicly available data and ultimately applying the gamma mixture model to this data.

## Description of data and data source
_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._

This study employs sales data, 'superstoredata' and focuses on international sales of the online retail store.

Initially this data was downloaded from Kaggle.com to it's usability for the project. Upon confirming the usability and revisiting the website seeking for further details, it could not be found in Kaggle.com. After few Google searches, the data and details were found on UC Irvine Machine Learning Repository [https://archive.ics.uci.edu/dataset/352/online+retail]. As per the website, this is a transnational data set containing all the transactions occurring between 12/1/2010 and 12/09/2011 for a UK-based and registered non-store online retail. The company primarily sells unique all-occasion gifts and many customers of the company are wholesalers.

This data has nine variables. The details are explained below:

(1)	InvoiceNo: It is a categorical variable with a 6-digit integral number uniquely assigned to each transaction. If this code starts with the letter ‘c’, it indicates a cancellation. 
(2)	StockCode: It is a categorical variable with a 5-digit integral number uniquely assigned to each distinct product.
(3)	Description: This represents Product name and is a categorical variable.
(4)	Quantity: It represents quantities of each product (item) per transaction and is an integer.
(5)	InvoiceDate: It represents the day and time each transaction was generated.
(6)	UnitPrice: It represents product price per unit in sterling.
(7)	CustomerID: It is a categorical variable with a 5-digit integral number uniquely assigned to each customer.
(8)	Country: It represents the name of the country where each customer resides.

The study will also use simulated data to formulate the EM algorithm. Following successful testing, the refined algorithm will be applied to the sales data.

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._

This study aims to answer the question, "How effective is the Expectation Maximization (EM) algorithm in identifying distinct customer groups when accounting for individual spending by each customer, as demonstrated by initial validation using simulated data?"

[1] Lee, S. X., & McLachlan, G. J. (2013). Model-based clustering and classification with non-normal mixture distributions. Statistical Methods & Applications, 22(4), 427-454.

[2] Young, D. S., Chen, X., Hewage, D. C., & Nilo-Poyanco, R. (2019). Finite mixture-of-gamma distributions: estimation, inference, and model-based clustering. Advances in Data Analysis and Classification, 13, 1053-1082.

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
#knitr::include_graphics(here("products","assets","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

This study employs sales data, 'superstoredata' and focuses on international sales of the online retail store.

Initially this data was downloaded from Kaggle.com to it's usability for the project. Upon confirming the usability and revisiting the website seeking for further details, it could not be found in Kaggle.com. After few Google searches, the data and details were found on UC Irvine Machine Learning Repository [https://archive.ics.uci.edu/dataset/352/online+retail]. As per the website, this is a transnational data set containing all the transactions occurring between 12/1/2010 and 12/09/2011 for a UK-based and registered non-store online retail. The company primarily sells unique all-occasion gifts and many customers of the company are wholesalers.

This data has nine variables. The details are explained below:

(1)	InvoiceNo: It is a categorical variable with a 6-digit integral number uniquely assigned to each transaction. If this code starts with the letter ‘c’, it indicates a cancellation. 
(2)	StockCode: It is a categorical variable with a 5-digit integral number uniquely assigned to each distinct product.
(3)	Description: This represents Product name and is a categorical variable.
(4)	Quantity: It represents quantities of each product (item) per transaction and is an integer.
(5)	InvoiceDate: It represents the day and time each transaction was generated.
(6)	UnitPrice: It represents product price per unit in sterling.
(7)	CustomerID: It is a categorical variable with a 5-digit integral number uniquely assigned to each customer.
(8)	Country: It represents the name of the country where each customer resides.

The study will also use simulated data to formulate the EM algorithm. Following successful testing, the refined algorithm will be applied to the sales data.

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "Sales_distribution.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","summarytable.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References