### 2024-03-07 R/eda-code/eda_superstore.qmd
Project Part 2 - Rev

---
title: "Exploratory Analysis of Superstore"
author: "Malika Dhakhwa"
date: "2024-02-23"
output: html_document
---

# Setup
Required packages are loaded.
```{r}

#| message: false
#| warning: false
#load needed packages. 
library(here) #for data loading/saving
library(dplyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(ggplot2) # for visualization
library(tibble) # For rownames_to_column
library(gt) # for table
library(gridExtra)
library(webshot2) #Capture as an Image the gttable before converting it to pdf
```

Data is loaded using here function.

```{r}
#Path to data using here() package
data_location <- here::here("data","processed-data","processed_superstore.rds")
#load data
mydata <- readRDS(data_location)
```

# Data exploration through tables

Creating summary table and saving it to the results folder.

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results","tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

# Descriptive Statistics tables

It is important to note that this data accounts sales of a particular item at a time as an observation leading different items sold in the same Invoice to appear as separate observations. As such most of the Invoice numbers appear several times in different observations of the data. A meaningful analysis of customer spending pattern require the Invoice Sales Amount, total purchase by a Customer during a period, frequency of purchase by a customer during a period or a combination of those. A table is created which provides the descriptive statistics of sales by Invoice value, Customer purchase and frequency of customer purchase. This table reveals that there were 1479 transactions to 414 different customers during 2010-12-01 and	2011-12-09.

```{r}

# Aggregate total sales by InvoiceNo for calculating mean, standard deviation, minimum and maximum
Sales_by_Invoice <- mydata %>%
  group_by(InvoiceNo) %>%
  summarise(TotalSales = sum(Sales, na.rm = TRUE)) %>%
  ungroup() %>% # Ensure we ungroup before further summarisation
  summarise(
    Count = n(),
    Particulars = "Sales by Invoice",
    Mean = round(mean(TotalSales, na.rm = TRUE), 2),
    SD = round(sd(TotalSales, na.rm = TRUE), 2),
    Min = round(min(TotalSales, na.rm = TRUE), 2),
    Max = round(max(TotalSales, na.rm = TRUE), 2)
  )

# Aggregate total sales by CustomerID to calculate mean, standard deviation, minimum and maximum
TotalSales_customer <- mydata %>%
  group_by(CustomerID) %>%
  summarise(TotalSales = sum(Sales, na.rm = TRUE)) %>%
  ungroup() %>%
  summarise(
    Count = n(),
    Particulars = "Cumulative Sales to Individual Customers",
    Mean = round(mean(TotalSales, na.rm = TRUE), 2),
    SD = round(sd(TotalSales, na.rm = TRUE), 2),
    Min = round(min(TotalSales, na.rm = TRUE), 2),
    Max = round(max(TotalSales, na.rm = TRUE), 2)
  )

# Frequency of sales per CustomerID (i.e., the count of unique InvoiceNo per CustomerID)for calculating mean, standard deviation, minimum and maximum

sales_frequency_per_customer <- mydata %>%
  group_by(CustomerID) %>%
  summarise(FrequencyOfSales = n_distinct(InvoiceNo)) %>%
  ungroup() %>%
  summarise(
    Count = n(),
    Mean = round(mean(FrequencyOfSales, na.rm = TRUE), 2),
    SD = round(sd(FrequencyOfSales, na.rm = TRUE), 2),
    Min = round(min(FrequencyOfSales, na.rm = TRUE), 2),
    Max = round(max(FrequencyOfSales, na.rm = TRUE), 2)
  ) %>%
  mutate(Particulars = "Sales frequency to Individual Customers")

# Combine the statistics into a single data frame
combined_stats <- bind_rows(Sales_by_Invoice, TotalSales_customer, sales_frequency_per_customer) %>%
  select(Particulars, everything())  # Ensure Particulars column is at the front

# Create and format the gt table
gt_table <- gt(combined_stats) %>%
  tab_header(
    title = "Descriptive Statistics of Sales") %>%
  cols_label(
    Particulars = "Particulars",
    Count = "Count",
    Mean = "Mean",
    SD = "St.Dev.",
    Min = "Min",
    Max = "Max"
  ) %>%
  cols_align(
    align = "center",
    columns = c("Count", "Mean", "SD", "Min", "Max")
  )

# Display the gt table
print(gt_table)

gt_table %>%
  gtsave(here("results","tables", "gt_table.png"))

```
# Data exploration through figures

Next, density plots were created for Sales across customers and Invoice Values. Both figures are heavily right-skewed.The averages of total sales to a customer and total invoice values appear to be less than 2500 and 1250 sterling pounds respectively supporting the mean values of 2025 and 563 in the descriptive table.

A histogram of purchase frequency of each customer is created by grouping each customer by invoice no.The average frequency appears to be less than 5 times for most of the customers which is also observed in the descriptive statistics table.

```{r}
# Aggregate sales by CustomerID
TotalSales_customer <- mydata %>%
  group_by(CustomerID) %>%
  summarise(TotalSales = sum(Sales, na.rm = TRUE)) %>%
  arrange(desc(TotalSales))

#Density Plot of Sales per Customer
# Assuming sales_per_customer is already calculated
p1 <- ggplot(TotalSales_customer, aes(x = TotalSales)) +
  geom_density(fill = "skyblue", alpha = 0.5) +
  labs(title = "Density Plot of Sales across Customers", x = "Total Sales", y = "Density")

# Histogram of Sales by Invoice values
## First calculate total invoice value per invoice
Sales_by_Invoice <- mydata %>%
  group_by(InvoiceNo) %>%
  summarise(TotalInvoiceValue = sum(Sales, na.rm = TRUE))

##Create histogram
p2 <- ggplot(Sales_by_Invoice, aes(x = TotalInvoiceValue)) +
  geom_density(fill = "lightgreen", alpha = 0.5) +
  labs(title = "Density Plot of  Invoice Sales",
       x = "Invoice Value", 
       y = "Density") +
  theme_minimal()

# Calculate purchase frequency per customer
purchase_frequency <- mydata %>%
  distinct(CustomerID, InvoiceNo) %>% # Ensure unique InvoiceNo per CustomerID
  count(CustomerID, name = "PurchaseFrequency") # Count the number of invoices per customer
library(ggplot2)

# Histogram of purchase frequencies
p3 <- ggplot(purchase_frequency, aes(x = PurchaseFrequency)) +
  geom_histogram(binwidth = 1, fill = "salmon", color = "darkblue") + # Adjust binwidth as needed
  labs(title = "Histogram of Customer Purchase Frequencies",
       x = "No. of Purchases", 
       y = "Count of Customers") +
  theme_minimal()

# Modify p1 to adjust title font size
p1 <- p1 + theme(plot.title = element_text(size = 11)) # Adjusting font size of the title of p1
# Modify p2 to adjust title font size
p2 <- p2 + theme(plot.title = element_text(size = 11)) # Adjusting font size of the title of p2
# Modify p3 to adjust title font size
p3 <- p3 + theme(plot.title = element_text(size = 11)) # Adjusting font size of the title of p3

plot1<-grid.arrange(p1, p2, p3, ncol = 2, nrow=2)

figure_file = here("results","figures","Sales_distribution.png")
ggsave(filename = figure_file, plot=plot1)

```

A histogram is also plotted for price range of the products.Most of the products fall below the price of 12.50 sterling pounds.

```{r}
# Create histogram of Unit Price

p4 <- ggplot(mydata, aes(x = UnitPrice)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "darkblue") + # Adjust binwidth as needed
  labs(title = "Histogram of Unit Price",
       x = "Unit Price", 
       y = "Frequency") +
  theme_minimal()

p4

figure_file = here("results","figures","Price_range.png")
ggsave(filename = figure_file, plot=p4)

```
Next, total purchase by a customer was plotted against purchase frequency.A few outliers were noted.

```{r}
#| message: false
#| warning: false
# Calculate purchase frequency and total invoice value per customer
customer_stats <- mydata %>%
  group_by(CustomerID) %>%
  summarise(PurchaseFrequency = n_distinct(InvoiceNo),
            TotalInvoiceValue = sum(Sales, na.rm = TRUE)) %>%
  ungroup()  # Ensure the data is ungrouped for further analysis

# Scatter plot of Total Invoice Value by Purchase Frequency
p5 <- ggplot(customer_stats, aes(x = PurchaseFrequency, y = TotalInvoiceValue)) +
  geom_point(alpha = 0.5) +  # Adjust alpha for point transparency, if needed
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear regression line without standard error
  labs(title = "Total Invoice Value by Purchase Frequency",
       x = "Purchase Frequency",
       y = "Total Invoice Value") +
  theme_minimal()
plot(p5)

figure_file = here("results","figures","Sales_trend.png")
ggsave(filename = figure_file, plot=p5)

```

Next, Bar chart is created for Sales Distribution by Country. The top five sales are in Germany, France, Spain, Switzerland and Japan.

```{r}
p8 <- mydata %>% 
  group_by(Country) %>% 
  summarise(TotalSales = sum(Sales)) %>%
  ggplot(aes(x = reorder(Country, TotalSales), y = TotalSales, fill = Country)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  theme(legend.position = "none")

plot(p8)

figure_file = here("results","figures","Countrywise_Sales.png")
ggsave(filename = figure_file, plot=p8)
```

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.